---
title: "Relief unit and raster objects"
author: "Gerald H. Taranto"
date: "`r format(Sys.time(), '%a %d, %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Relief unit and raster objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(out.width = "100%", echo = TRUE)
```

## Load data

Libraries and data are loaded and processed as explained in [format input data](scapesClassification_02_1_DATA.html). 

```{r loadLD, message=FALSE, warning=FALSE, collapse = TRUE, class.source = 'fold-hide'}
# LOAD LIBRARIES
library(scapesClassification)
library(terra)

# LOAD DATA
grd <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T)
grd <- grd[grepl("\\.grd", grd)]
grd <- grd[!grepl("hillshade", grd)]
rstack <- rast(grd)

# COMPUTE ATTRIBUTE TABLE
atbl <- attTbl(rstack, var_names = c("bathymetry", "local_bpi", "regional_bpi", "slope"))

# COMPUTE NEIGHBORHOOD LIST
nbs <- ngbList(rstack, rNumb = TRUE, attTbl = atbl) # neighbors are identified by their row number in the attribute table
```

### Class vector filters

We will also add to the attribute table the class vector of the [island shelf unit (ISU)](scapesClassification_02_2_ISU.html) and the class vector of the [peak cells](scapesClassification_02_3_PKS.html) so that they can be used to build classification rules. 

```{r loadCV, message=FALSE, warning=FALSE}
# LOAD ISU CLASSIFICATION
ISU <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T,
                  pattern = "ISU_Cells.RDS")

atbl$ISU <- readRDS(ISU)
atbl$ISU[!is.na(atbl$ISU)] <- 1 # Class 1 identifies all ISU cells 

# LOAD PEAK CLASSIFICATION
PKS <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T, 
                  pattern = "PKS_Cells.RDS")
atbl$PKS <- readRDS(PKS)

head(atbl)
```

### Plots

In the following example we will show how the class vectors of raster objects are computed. However, in order to improve the reading experience, the plots' code is hidden. It can be accessed in the `*.RMD` file used to generate the html file. 

The plotting procedure is to (i) convert a class vectors into a raster using the function `cv.2.rast()` and to (ii) visualize the raster using R or an external software. In our example we will use the R package `terra` to create static maps and the R packages `mapview` and `leaflet` to create interactive maps (_note that `mapview` do not support `terra` raster objects yet, therefore, they have to be converted into `raster` raster objects before plotting_). 

## Relief unit

The relief unit (RU) is composed of prominent seamounts, banks and ridges. In this article individual elements of the relief unit will be classified as distinct raster objects each characterized by a unique ID ([Figure 1](#figure1) and [Figure 4](#figure4)). The set of all objects will define the relief unit of our working example.

### RU objects

In this example, relief units objects will be identified in a sequence of classification steps using [raster object functions](https://ghtaranto.github.io/scapesClassification/reference/index.html#manipulation-functions) ([Figure 1](#figure1)). Our functions will perform the following tasks: 

1. **[anchor.seed](scapesClassification_01_4_ObjFunctions.html#anchor-seed):** identify all raster cells (i) on prominent relieves that are (ii) not ISU cells and that are (iii) in continuity with peak cells. It assigns a different ID to each non-continuous group of cell ([Figure 1a](#figure1)).

2. **[rel.pi](scapesClassification_01_4_ObjFunctions.html#rel-pi):** compute the standardized relative position index (`rPI`) of each raster object. 

3. **[pi.sgm](scapesClassification_01_4_ObjFunctions.html#pi-sgm):** (i) segment raster objects at locations that have negative `rPI` or [local bpi](scapesClassification_02_1_DATA.html#study-case-and-input-data) values; (ii) remove small raster objects composed of less than 40 cells ([Figure 1b](#figure1)). 

4. **[pi.add](scapesClassification_01_4_ObjFunctions.html#pi-add):** add new raster objects composed of (i) cells that have [local bpi](scapesClassification_02_1_DATA.html#study-case-and-input-data) values `>100` **and** (ii) that do not border any other object (i.e. an object is added only if it is disjoint from other objects) ([Figure 1c](#figure1)). 

5. **[cond.4.nofn](scapesClassification_01_4_ObjFunctions.html#hgrowth):** homogeneous growth of raster objects. At turns, each object adds contiguous cells having [local or regional bpi](scapesClassification_02_1_DATA.html#study-case-and-input-data) values `>100` ([Figure 1d](#figure1)). 

--- 

_Additional information about the above functions is available at the following links: [function documentation](https://ghtaranto.github.io/scapesClassification/reference/index.html#manipulation-functions) and [object functions](scapesClassification_01_4_ObjFunctions.html) (article)._

---

```{r ro_cell, message=FALSE, warning=FALSE, collapse = TRUE}
# 1. ANCHOR.SEED
# Tasks:
# Identify raster cells on prominent features in connection with peak cells
# that are not within island shelf units (ISU).
atbl$RO <- anchor.seed(atbl, nbs, rNumb = TRUE, silent = TRUE,
                       
                       class = NULL, # A new ID is assigned to every
                                     # discrete group of cells...
                      
                       cond.seed   = "PKS == 1", # ...in continuity with peak cells...
                       
                       cond.growth = "regional_bpi >= 100", # ...that are on prominent features.
                       
                       cond.filter = "regional_bpi >= 100 & is.na(ISU)") # Only consider non-ISU cells 
                                                                         # on prominent features.  

rRO1 <- cv.2.rast(rstack, atbl$RO) # store raster for plotting
#######################################################################################################

# 2. REL.PI
# Tasks: 
# Compute the standardized relative position index (`rPI`) of each raster object.
atbl$rPI <- rel.pi(atbl, RO="RO", el="bathymetry")
#######################################################################################################

# 3. PI.SGM
# Tasks:
# (i) Segment large raster objects that have negative rPI or negative local bpi;
# (ii) Remove raster objects with less than 40 cells.
atbl$RO <- pi.sgm(atbl, nbs, rNumb = TRUE,
                  RO = "RO", 
                  mainPI = "rPI",
                  secPI = "local_bpi", 
                  cut.mPI = 0, 
                  cut.sPI = 0,
                  min.N = 40)

rRO2 <- cv.2.rast(rstack, atbl$RO) # store raster for plotting
#######################################################################################################

# 4. PI.ADD
# Tasks:
# Add new raster objects composed of cells with local_bpi>100; 
# An object is added only if it is disjoint from other objects,
# if it has more than 40 cells and if it is not within the ISU.
atbl$RO <- pi.add(atbl, nbs, rNumb = TRUE,
                  RO = "RO",
                  mainPI = "local_bpi", 
                  add.mPI = 100, 
                  min.N = 40,
                  cond.filter = "is.na(ISU)")

rRO3 <- cv.2.rast(rstack, atbl$RO) # store raster for plotting
#######################################################################################################

# 5. COND.4.NOFN (HGROWTH)
# Task:
# Homogeneous growth of raster objects;
# At turns, each raster object adds contiguous cells having 
# regional or local BPI values > 100
IDs <- unique(atbl$RO)[!is.na(unique(atbl$RO))]

atbl$RO <- cond.4.nofn(atbl, nbs, rNumb = TRUE, classVector = atbl$RO,
                       nbs_of = IDs, class = NULL,
                       cond = "regional_bpi > 100 | local_bpi > 100",
                       hgrowth = TRUE) 

rRO4 <- cv.2.rast(rstack, atbl$RO) # store raster for plotting
```

<a id="figure1"></a>

```{r ro_plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=7}
# COMPUTE BORDERS FOR PLOTTING
RO_bd <- obj.border(split(1:NROW(atbl), atbl$RO), nbs, silent = TRUE)
RO_bd <- obj.nbs(RO_bd, nbs, silent = TRUE)
RO_bd <- unlist(unique(RO_bd))
RO_bd <- atbl$Cell[RO_bd]

rRO_bd <- rRO3; rRO_bd[] <- NA; rRO_bd[ RO_bd ] <- 1


layout(matrix(c(1, 2,
                1, 2,
                3, 4, 
                3, 4), nrow=4, byrow=TRUE))

m <- c(1.5,1,2,1)

xl <- ext(rRO1)[c(1,2)]
yl <- ext(rRO1)[c(3,4)]

# PLOT 1
plot(rRO1, type = "classes", legend= FALSE, mar=m, xlim=xl, ylim=yl)
title("(a) anchor.seed", adj = 0, family = "serif", cex = 1.2)

# PLOT 2
plot(rRO2, type = "classes", legend= FALSE, mar=m, xlim=xl, ylim=yl)
title("(b) pi.sgm", adj = 0, family = "serif", cex = 1.2)

# PLOT 3
plot(rRO3, type = "classes", legend= FALSE, mar=m, xlim=xl, ylim=yl)
title("(c) pi.add", adj = 0, family = "serif", cex = 1.2)

# PLOT 4
plot(rRO4, type = "classes", legend= FALSE, mar=m, xlim=xl, ylim=yl)
title("(d) cond.4.nofn (hgrowth)", adj = 0, family = "serif", cex = 1.2)
plot(rRO_bd, col = "black", legend= FALSE, add = TRUE)
```

<div style="line-height: 0.9em; margin-bottom: 1.5em;"><span class="legend">**Figure 1 - Relief unit raster objects,** colors identify discrete raster objects. Figure panels show the classification steps used to identify the different raster objects. In panel (d) black lines show the borders of contiguous objects.</span></div>

### Relative position index

As we saw in the previous section, relative position indices are useful to segment and to identify raster objects. Another important quality of position indices is that they allow to visualize the depth distribution of individual objects and, thus, discern where summits and bases are ([Figure 2](#figure2)).

```{r rPI, message=FALSE, warning=FALSE, collapse = TRUE}
# NORMALIZED RELATIVE POSITION INDEX
atbl$rPI <- rel.pi(atbl, RO="RO", el="bathymetry", type = "n")
```

<a id="figure2"></a>

```{r rPI_plot, message=FALSE, warning=FALSE, echo=FALSE}
# PEAK CELLS
library(data.table)
atbl <- as.data.table(atbl)
atbl[!is.na(RO), mbat := bathymetry == max(bathymetry), by = "RO"]
pc <- atbl[which(mbat), Cell]
atbl$mbat <- NULL

# PLOT
r_rPI <- cv.2.rast(rstack, atbl$rPI)
brk <- seq(0,1,0.2)
terra::plot(r_rPI, type="interval", breaks=brk, 
            plg = list(title="rPI", title.adj=0.1))
terra::plot(rRO_bd, col = "black", legend= FALSE, add = TRUE)

terra::points(xyFromCell(rstack, pc), pch=20, col="red", cex=0.9)

legend(ext(r_rPI)[2], ext(r_rPI)[4]-25000, legend=c("Peak"), bty = "n", xpd=TRUE,
       col=c("red"), pch=20, cex=0.9, pt.cex = 1.5)
```

<div style="line-height: 0.9em; margin-bottom: 1.5em;"><span class="legend">**Figure 2 - Normalized relative position index,** normalized depth values computed independently for each raster objects. High values point to summit areas, low values to bottom areas. Black lines show the borders of contiguous objects</span></div>

### Ridges

Ridge systems are an ecologically important component of relieves. In this section we will identify the distinct ridges existing within the relief unit. 
We will adopt a simple definition of what a ridge is: 

* They are within raster objects;

* They have high [local bpi](scapesClassification_02_1_DATA.html#study-case-and-input-data) values (`>100`);

* They have a minimum of 20 cells.

The first two rules are implemented using [anchor.seed](scapesClassification_01_4_ObjFunctions.html#anchor-seed), the last using [pi.sgm](scapesClassification_01_4_ObjFunctions.html#pi-sgm) ([Figure 3](#figure3)).

```{r ridge_cell, message=FALSE, warning=FALSE, collapse = TRUE}
# IDENTIFY RIDGES
atbl$RDG <- anchor.seed(atbl, nbs, rNumb = TRUE, silent = TRUE, class = NULL,
                        cond.seed   = "local_bpi > 100",
                        cond.growth = "local_bpi > 100",
                        cond.filter = "!is.na(RO)")

# REMOVE SMALL RIDGES
atbl$RDG <- pi.sgm(atbl, nbs, rNumb = TRUE, RO = "RDG", min.N = 20)
```

<a id="figure3"></a>

```{r ridge_plot, message=FALSE, warning=FALSE, echo=FALSE}
# ALL ROs HAVE THE SAMWE CLASS
ROall <- atbl$RO
ROall[!is.na(ROall)] <- 1
rROall <- cv.2.rast(rstack, ROall)

# RIDGES
rRDG <- cv.2.rast(rstack, atbl$RDG)

plot(rROall, type="classes", col = "black", ext = ext(rROall),
     plg=list(x=ext(rROall)[2], y=ext(rROall)[4], legend = "RU"))
plot(rRDG, type="classes", ext = ext(rROall), add = TRUE,
     plg=list(x=ext(rROall)[2], y=ext(rROall)[4]-10000, title = "Ridges (#14):", ncol=2))
```

<div style="line-height: 0.9em; margin-bottom: 1.5em;"><span class="legend">**Figure 3 - Ridges,** colors identify discrete ridge objects. In black is highlighted the relief unit (RU).</span></div>

## Data.table

The R package [data.table](https://rdatatable.gitlab.io/data.table/) provides an easy syntax to handle and transform groups within tables and facilitates the computation of group statistics. In the context of `scapesClassification`, it can be used to facilitate hierarchic classifications and to summarize raster object statistics. 

As an example, we will first compute some simple statistics for each raster object: number of cells (`N`), average (`avgD`), maximum (`maxD`), minimum (`minD`) and standard deviation (`sdD`) of depth values. In addition, we will count the number of discrete ridges each object contains (`nRDG`) and the percentage of the object they cover (`pRDG`).  

Finally we will rename raster object IDs based on the rank of minimum depth values: the raster object with the lowest minimum depth value will have `ID = 1`; the object with the highest minimum depth value will have the highest ID. 

```{r data_table, message=FALSE, warning=FALSE}
# DATA TABLE
library(data.table)

# CONVERT ATTRIBUTE TABLE TO ATTRIBUTE 'DATA' TABLE
atbl <- as.data.table(atbl)

# COMPUTE STATISTICS
RO_stat <- atbl[!is.na(RO), 
                
                .(N = .N,
                  minD = round(max(bathymetry), 2),
                  avgD = round(mean(bathymetry), 2),
                  maxD = round(min(bathymetry), 2),
                  sdD = round(sd(bathymetry), 2), 
                  nRDG = sum(!duplicated(RDG), na.rm = TRUE),
                  pRDG = round(sum(!is.na(RDG))/.N, 2)),
                
                by = "RO"]

# RENAME AND RANK RO IDs BY MIN. DEPTH
new_IDs <- RO_stat[, data.table::frankv(minD, order = -1L)]

atbl[!is.na(RO), RO := new_IDs[match(RO, RO_stat$RO)] ]
RO_stat[, RO := new_IDs][order(RO)][]
```

## Interactive maps

Raster objects can be visualized as interactive maps using the R packages [mapview](https://r-spatial.github.io/mapview/) and [leaflet](https://rstudio.github.io/leaflet/). Each object can be associated to a pop-up using the R package [leafpop](https://github.com/r-spatial/leafpop). 
The combined use of `scapesClassification`, `data.table` and interactive maps provides users with the capacity to highlight areas of interest, summarize raster object statistics and share the outcome. 

As an example we will produce an interactive map of the relief unit objects with pop-up windows presenting the statistics computed with `data.table` ([Figure 4](#figure4)). 

```{r loadMapView, message=FALSE, include=FALSE}
library(mapview)
library(leaflet)
library(leafem)
library(leafpop)

# set gerenal options 
mapviewOptions(basemaps = c("Esri.OceanBasemap", "OpenStreetMap"), na.color = "grey88", homebutton = FALSE, 
               query.position = "bottomright", legend.opacity = 0.9)

# set palettes
palRYB <- c("#2892c7", "#fafa64", "#e81014")
palYGB <- c("#ffff80" ,"#3dba65", "#0d1178")
palBYR <- c("#4575b5", "#ffffbf", "#d62f27")
gm_col <- c("grey70", "#380000", "#c80000", "#ff5014", "#fad23c", "#ffff3c", "#b4e614", "#3cfa96", "#0000ff", "#000038")
```

<a id="figure4"></a>

```{r RO_mapview, message=FALSE, warning=FALSE, echo=FALSE, fig.height=6}
# RASTER OBJECTS
rRO <- cv.2.rast(rstack, atbl$RO)
rRO <- as.polygons(rRO, dissolve = TRUE)

# RIDGES
RDG  <- atbl$RDG
RDG[which(!is.na(RDG))] <- 1
rRDG <- cv.2.rast(rstack, RDG)
rRDG <- as.polygons(rRDG, dissolve = TRUE)

# PEAKS
pcV <- vect(xyFromCell(rstack, pc), crs = crs(rRO))

# ISU
rISU <- cv.2.rast(rstack, atbl$ISU)
rISU <- as.polygons(rISU, dissolve = TRUE)

# BUILD POP UP TABLE
names(rRO) <- "ID"
names(rRDG)<- "ID"

# Index vector IDs into RO_stat
ind <- match(rRO$ID, RO_stat$RO)

# Add columns
rRO$N <- RO_stat$N[ind]
rRO$minD <- RO_stat$minD[ind]
rRO$avgD <- RO_stat$avgD[ind]
rRO$maxD <- RO_stat$maxD[ind]
rRO$sdD <- RO_stat$sdD[ind]
rRO$nRDG <- RO_stat$nRDG[ind]
rRO$pRDG <- RO_stat$pRDG[ind]

# Plot
m<- mapview(as(rRO,"Spatial"), layer.name = "RO", zcol = "ID", alpha.regions = 0.8, legend = FALSE,
            popup = popupTable(as(rRO,"Spatial"), feature.id = FALSE, row.numbers = FALSE)) +
    mapview(as(rRDG,"Spatial"), layer.name = "RDG", alpha.regions = 0.6, col.regions = "white", legend= FALSE, 
          popup = FALSE) +
    mapview(as(rISU,"Spatial"), layer.name = "ISU", alpha.regions = 0.6, col.regions = "black", legend= FALSE, 
          popup = FALSE) +
    mapview(as(pcV,"Spatial"), layer.name = "Peak", alpha.regions = 0.6, col.regions = "red", legend= FALSE,
          popup = FALSE) + 
    mapview(raster::raster(rstack[["bathymetry_exm"]]), layer.name = "Depth (m)", col.regions = palRYB) +
    mapview(raster::raster(rstack[["local_bpi_exm"]]), layer.name = "Local BPI", col.regions = palBYR)

m@map %>% 
  hideGroup(c("Depth (m)", "RDG", "Local BPI", "Peak", "ISU")) %>% 
  addLayersControl(baseGroups = c("OpenStreetMap", "Esri.OceanBasemap"),
                   overlayGroups = c("RO", "RDG", "Peak", "ISU", "Depth (m)", "Local BPI"),
                   position = "topleft",
                   options = layersControlOptions(collapsed = TRUE))
```

<div style="line-height: 0.9em; margin-bottom: 1.5em;"><span class="legend">**Figure 4 - Raster object interactive map,** colors identify discrete raster objects. **RO**, raster objects (clickable layer). **RDG**, ridges within raster objects. **Peak**, main peak of each raster object. **ISU**, island shelf unit. Bathymetry and local BPI values are provided in background (due to plot re-projection, displayed and original values might be slightly different).</span></div>

## Final class vector

The relief unit class vector is saved in the file `"RU_obj.RDS"` and shared as part of the package data. 

```{r classVect_RO, message=FALSE, warning=FALSE, collapse = TRUE}
# RASTER OBJECTS
unique(atbl$RO)

# A total of 12 raster objects were identified
length(na.omit(unique(atbl$RO)))
```
