---
title: "Geomorphic Management Units (GMUs)"
author: "Gerald H. Taranto"
date: "`r format(Sys.time(), '%a %d, %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Geomorphic Management Units (GMUs)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(out.width = "100%", echo = TRUE)
```

The following articles explain how to use how to use [scapesClassification](https://github.com/ghTaranto/scapesClassification) to classify marine seafloor structures, hereafter referred to as geomorphic management units (GMUs). 

The GMU-classification will be presented in four separate articles:

1. [Format input data](scapesClassification_02_1_DATA.html); 

2. [Island shelf unit (ISU)](scapesClassification_02_2_ISU.html)

3. [Local maxima on gridded data](scapesClassification_02_3_PKS.html); 

4. [Independent objects on raster surfaces](scapesClassification_02_4_OBJ.html). 



```{r loadLD, message=FALSE, warning=FALSE, include=FALSE}
# LOAD LIBRARIES
library(raster)
library(scapesClassification)

# LOAD DATA
grd <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T)
grd <- grd[grepl("\\.grd", grd)]
grd <- grd[!grepl("hillshade", grd)]
rstack <- stack(grd)

# COMPUTE ATTRIBUTE TABLE
atbl <- attTbl(rstack, var_names = c("bathymetry", "gmorph", "local_bpi", "regional_bpi", "sd_bat", "slope"))

# COMPUTE NEIGHBORHOOD LIST
nbs <- ngbList(rstack, rNumb = TRUE, attTbl = atbl) # neighbors are identified by their row number in the attribute table

# LOAD ISU CLASSIFICATION
ISU <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T, 
                  pattern = "ISU_Cells\\.RDS")

ISU <- readRDS(ISU)

# LOAD PEAK CLASSIFICATION
PKS <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T, 
                  pattern = "Peak_Cells\\.RDS")

PKS <- readRDS(PKS)

# LOAD RU_OBJECT CLASSIFICATION
RU_OBJ <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T, 
                  pattern = "RU_obj\\.RDS")

RU_OBJ <- readRDS(RU_OBJ)

RU_OBJ2 <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T, 
                  pattern = "RU_obj2\\.RDS")

RU_OBJ2 <- readRDS(RU_OBJ2)


# LOAD PLOT LIBRARIES
library(mapview)
library(leaflet)
library(leafem)

# set gerenal options 
mapviewOptions(basemaps = c("Esri.OceanBasemap", "OpenStreetMap"), na.color = "grey88", homebutton = FALSE, 
               query.position = "bottomright", legend.opacity = 0.9)

# hillShade
hs <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T)
hs <- raster(hs[grepl("hillshade\\.grd", hs)])

# set palettes
palRYB <- c("#2892c7", "#fafa64", "#e81014")
palYGB <- c("#ffff80" ,"#3dba65", "#0d1178")
palBYR <- c("#4575b5", "#ffffbf", "#d62f27")
gm_col <- c("grey70", "#380000", "#c80000", "#ff5014", "#fad23c", "#ffff3c", "#b4e614", "#3cfa96", "#0000ff", "#000038")
```


# Island shelf unit (ISU)
```{r ISU_slope_map, message=FALSE, warning=FALSE, echo=FALSE}
# ISU
ISU_shelf <- ISU
ISU_shelf[ISU_shelf %in% 1:3] <- 1
ISU_shelf[ISU_shelf %in% 4:6] <- NA

ISU_slope <- ISU
ISU_slope[ISU_slope %in% 1:3] <- NA
ISU_slope[ISU_slope %in% 4:6] <- 1

ISU_shelf <- cv.2.rast(rstack,atbl$Cell, ISU_shelf)
ISU_shelf <- rasterToPolygons(ISU_shelf, dissolve = TRUE)

ISU_slope <- cv.2.rast(rstack,atbl$Cell, ISU_slope)
ISU_slope <- rasterToPolygons(ISU_slope, dissolve = TRUE)

ISU_shelf$dummy <- 1
ISU_slope$dummy <- 1


m <-
  mapview(hs, col.regions = c("gray0", "gray50", "grey100"), alpha.regions=1, layer.name = "hillshade", legend = FALSE, label = FALSE) + 
  
  mapview(rstack[["bathymetry_exm"]], layer.name = "bathymetry(m)", col.regions = palRYB, legend = FALSE, alpha.regions = 0.5) +
  
  mapview(ISU_shelf, alpha.regions = 1, col.regions = "#02343F", layer.name = "ISU-Shelf", popup = NULL, 
          label = "ISU-Shelf") +
  
  mapview(ISU_slope, alpha.regions = 0.8, layer.name = "ISU-Slope", col.regions = "#F0EDCC", popup = NULL, 
          label = "ISU-Slope")
  

m@map %>% 
  hideGroup(c("hillshade", "bathymetry(m)")) %>% 
  addLayersControl(overlayGroups = c("hillshade", "bathymetry(m)", "ISU-Shelf", "ISU-Slope"),
                   position = "topleft",
                   options = layersControlOptions(collapsed = FALSE))
```
Figure 1 - Island shelf unit. Computed in [island shelf unit (ISU)](scapesClassification_02_2_ISU.html). 


# Local peaks
```{r peaks_map, message=FALSE, warning=FALSE, echo=FALSE}
pks_plot <- cv.2.rast(rstack, atbl$Cell, PKS)
pks_plot <- rasterToPolygons(pks_plot, dissolve = TRUE)
pks_plot$dummy <- 1

m <- mapview(hs, col.regions = c("gray0", "gray50", "grey100"), alpha.regions=1, 
             legend = FALSE, label = FALSE) + 
  mapview(rstack[["bathymetry_exm"]], layer.name = "Bathymetry(m)", col.regions = palRYB, 
          legend = FALSE, alpha.regions = 0.5) + 
  mapview(pks_plot, layer.name = "Local peaks", col.regions = "black", label = "Local peaks", 
          alpha.regions = 1,  popup = NULL)

m@map %>% 
  addLayersControl(overlayGroups = c("Local peaks"),
                   position = "topleft",
                   options = layersControlOptions(collapsed = FALSE))
```
Figure 2 - Local peaks. Computed in [local maxima on gridded data](scapesClassification_02_3_PKS.html).

# RU-objects
```{r RU_obj, message=FALSE, warning=FALSE, echo=FALSE}
library(data.table)
library(leafpop)

atbl <- as.data.table(atbl)

atbl$RU_OBJ  <- RU_OBJ
atbl$RU_OBJ2 <- RU_OBJ2

# POPUP TABLES
pupTbl <- atbl

pupTbl[!is.na(RU_OBJ),
       
       ':=' (ncell = .N,
             minDepth_m = round(max(bathymetry), 2),
             maxDepth_m = round(min(bathymetry), 2),
             avgDepth_m = round(mean(bathymetry), 2),
             area_m2 = 500 * .N),
       
       by = "RU_OBJ"]

pupTbl[!is.na(RU_OBJ2),
       
       ':=' (ncell_2 = .N,
             minDepth_m_2 = round(max(bathymetry), 2),
             maxDepth_m_2 = round(min(bathymetry), 2),
             avgDepth_m_2 = round(mean(bathymetry), 2),
             area_m2_2 = 500 * .N),
       
       by = "RU_OBJ2"]


pupTbl_stat  <- pupTbl[!duplicated(RU_OBJ) & !is.na(RU_OBJ)]

# RU-objects
RU_obj_plot <- cv.2.rast(rstack, atbl$Cell, RU_OBJ)
RU_obj_plot <- rasterToPolygons(RU_obj_plot, dissolve = TRUE)

RU_obj2_plot <- cv.2.rast(rstack, atbl$Cell, RU_OBJ2)
RU_obj2_plot <- rasterToPolygons(RU_obj2_plot, dissolve = TRUE)


# POLYGON TABLES
names(RU_obj_plot)  <- "RU_obj_ID"
names(RU_obj2_plot) <- "RU_obj_ID"

ind <- match(RU_obj_plot@data$RU_obj_ID, pupTbl_stat$RU_OBJ)

RU_obj_plot@data$ncell      <- pupTbl_stat$ncell[ind]
RU_obj_plot@data$minDepth_m <- pupTbl_stat$minDepth_m[ind]
RU_obj_plot@data$maxDepth_m <- pupTbl_stat$maxDepth_m[ind]
RU_obj_plot@data$avgDepth_m <- pupTbl_stat$avgDepth_m[ind]
RU_obj_plot@data$area_m2    <- pupTbl_stat$area_m2[ind]

RU_obj2_plot@data$ncell      <- pupTbl_stat$ncell_2[ind]
RU_obj2_plot@data$minDepth_m <- pupTbl_stat$minDepth_m_2[ind]
RU_obj2_plot@data$maxDepth_m <- pupTbl_stat$maxDepth_m_2[ind]
RU_obj2_plot@data$avgDepth_m <- pupTbl_stat$avgDepth_m_2[ind]
RU_obj2_plot@data$area_m2    <- pupTbl_stat$area_m2_2[ind]


# plot
m <- mapview(hs, col.regions = c("gray0", "gray50", "grey100"), alpha.regions=1, legend = FALSE, label = FALSE) +
  
  mapview(rstack[["bathymetry_exm"]], layer.name = "bathymetry", col.regions = palRYB, legend = FALSE, alpha.regions = 0.5) +

  mapview(RU_obj_plot, layer.name = "RU-object, core", alpha.regions = 0.8, legend = FALSE, zcol = "RU_obj_ID",
          popup = popupTable(RU_obj_plot, feature.id = FALSE, row.numbers = FALSE)) + 
  
  mapview(RU_obj2_plot, layer.name = "RU-object, all", alpha.regions = 0.8, legend = FALSE, zcol = "RU_obj_ID",
          popup = popupTable(RU_obj2_plot, feature.id = FALSE, row.numbers = FALSE)) + 
  
  mapview(pks_plot, layer.name = "Local peaks", col.regions = "black", label = "Local peaks", 
          alpha.regions = 1,  popup = NULL)

m@map %>%
  hideGroup(c("RU-object, all", "Local peaks")) %>%
  addLayersControl(overlayGroups = c("Local peaks", "RU-object, core", "RU-object, all"),
                   position = "topleft",
                   options = layersControlOptions(collapsed = FALSE))
```
Figure 3 - Relief unit objects, clickable layers. Computed in [independent objects on raster surfaces](scapesClassification_02_4_OBJ.html). 
