---
title: "Classification functions"
author: "Gerald H. Taranto"
date: "`r format(Sys.time(), '%a %d, %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Classification functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', message=FALSE, warning=FALSE, out.width = "100%")
```

## Class vectors

Class vectors map raster cells to numeric classes (or to NA-values if no class is attributed). The n^th^ element of a class vector corresponds to the raster cell stored in the n^th^ row of an [attribute table](scapesClassification_01_1_FormatInputs.html#attribute-table). Class vectors are used by classification functions to store the results of their classifications. 

Class vectors can also be used as function inputs. As inputs, they provide information about what cells have been classified and allow to consider class contiguity and class continuity in rule evaluation (see the article ["rule evaluation"](scapesClassification_01_2_RuleEvaluation.html)). 

When a class vector is provided as an input, classification functions assign a class only to unclassified cells (i.e. cells corresponding to NA values in the class vector). In this way class vectors are updated (and not overwritten) every time the are provided as a function input. The function `anchor.seed()` is the only exception. 

## Anchor cell functions

**Functions:** `anchor.svo()`, `anchor.cell()` and `cond.4.all()`.

Anchor cells can be thought of as raster cells having particular characteristics from where the classification process can start. Anchor cells can be derived by: previous classifications, areas of constant environmental or geographic characteristics (e.g., flat areas), local maxima or minima (e.g., local peaks on a mountain ridge) or spatial objects (e.g., shapefiles).  

Anchor cell functions can be easily identified because they do not necessarily require a class vector as an input. 

### anchor.svo and anchor.cell

To illustrate the functions `anchor.svo()` and `anchor.cell()` we will use the data stored within the package. These data include several raster layers around the central group of the Azores Islands and the shapefile of the islands (see the article ["island shelf unit"](scapesClassification_02_2_ISU.html) for additional details).  

We will start by loading the data and the libraries and by computing the attribute table and the list of neighborhoods. 

```{r loadLD, message=FALSE, warning=FALSE, collapse = TRUE, class.source = 'fold-hide'}
# LOAD LIBRARIES
library(raster)
library(scapesClassification)

# LOAD DATA
grd <- list.files(system.file("extdata", package = "scapesClassification"), full.names = T)
grd <- grd[grepl("\\.grd", grd)]
grd <- grd[!grepl("hillshade", grd)]
rstack <- stack(grd)

# COMPUTE ATTRIBUTE TABLE
atbl <- attTbl(rstack, var_names = c("bathymetry", "gmorph", "local_bpi", "regional_bpi", "sd_bat", "slope"))

# COMPUTE NEIGHBORHOOD LIST
nbs <- ngbList(rstack, rNumb = TRUE, attTbl = atbl) # neighbors are identified by their row number in the attribute table

# ISLAND SHAPEFILE PATH
shp <- system.file("extdata", "Azores.shp", package = "scapesClassification")
```

The function `anchor.svo()` returns a vector of cell numbers extracted at the locations of a spatial vector object. In this case it will extract "land cells" at the locations of the island shapefile. 

Function arguments:

* `only_NAs = TRUE`, cell numbers extracted only at the locations of a spatial vector object that are not complete cases (i.e. have some missing values in some of the layers of the `Raster*` object).

* `fill_NAs = TRUE`, cell numbers extracted also at locations contiguous to those of the spatial vector object that are not complete cases (i.e. have some missing values in some of the layers of the `Raster*` object). 

```{r ISU_land, message=FALSE, warning=FALSE, collapse = TRUE}
# EXTRACT LAND POSITIONS
anchorLAND <- anchor.svo(rstack              = rstack, 
                         spatial_vector_name = shp, 
                         only_NAs            = TRUE, 
                         fill_NAs            = TRUE, 
                         plot                = FALSE)

# CELL NUMBERS AT THE LOCATION OF `shp`
head(anchorLAND)
```

The function `anchor.cell()` converts a vector of cell numbers into a class vector. In this case we want to attribute a classification number (argument `class = 1`) to cells adjacent to land cells.

Function arguments:

* `class2cell = FALSE`, the classification number is not attributed to cells included in the vector `anchorLAND`. 

* `class2nbs = TRUE`, the classification number is attributed to cells adjacent to those included in the vector `anchorLAND`. 

```{r ISU_anchor, message=FALSE, warning=FALSE, collapse = TRUE}
# COMPUTE ANCHOR CELLS
anchorCELL <- anchor.cell(attTbl     = atbl, 
                          rstack     = rstack, 
                          anchor     = anchorLAND, 
                          class      = 1,
                          class2cell = FALSE,
                          class2nbs  = TRUE,
                          plot       = FALSE) 

# EXTRACT anchorCELL FROM THE ATTRIBUTE TABLE
head( atbl$Cell[which(anchorCELL == 1)] )
```

Finally we can visualize our results using the packages `mapview` and `leaflet` ([Figure 1](#figure1)).

<a id="figure1"></a>
```{r leaflet, message=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
library(mapview)
library(leaflet)

# set gerenal options 
mapviewOptions(basemaps = c("OpenStreetMap"), na.color = "grey88", homebutton = FALSE, 
               query.position = "bottomright", legend.opacity = 0.9)

# ISU_anchor
island <- rgdal::readOGR(dsn = shp, verbose = F)
island <- sp::spTransform(island, raster::crs(rstack))
island$BASE[c(5,4,49)] <- c("Pico", "Faial", "Sao Jorge") # mouseover labels

rL  <- rstack[[1]]; rL[]  <- NA; rL[anchorLAND] <- 1
rAC <- rstack[[1]]; rAC[] <- NA; rAC[atbl$Cell] <- anchorCELL

# convert to polygon so that mapview projection does not affect visualization
rL  <- rasterToPolygons(rL, dissolve = TRUE)
rAC <- rasterToPolygons(rAC, dissolve = TRUE)

# add a column so that mapview legend reads layer.name and not values
rL$dummy  <- 1
rAC$dummy <- 1

mv <- 
  mapview(rAC, alpha.regions = 1, 
             col.regions = "orange", layer.name = "Anchor cell", popup = NULL, label = "Anchor Cell") + 
  mapview(rL, alpha.regions = 1, 
          col.regions = "black", layer.name = "Land cell", popup = NULL, label = "Land cell") + 
  mapview(island, alpha.region = 0, col.regions = "cyan", color = "cyan", lwd = 2, label = "BASE", 
          layer.name = "Island shapefile", popup = NULL)

mv@map %>% 
  setView(lng = -28.35, lat = 38.59, zoom = 10) %>%
  addLayersControl(baseGroups = c("OpenStreetMap", "Esri.OceanBasemap"),
                   overlayGroups = c("Anchor cell", "Land cell", "Island shapefile"),
                   position = "topleft",
                   options = layersControlOptions(collapsed = TRUE))
```

<div style="line-height: 0.8em; margin-bottom: 1.5em;"><span class="legend">**Figure 1 -** **Anchor cell:** raster cells adjacent to a 'land cell'. **Land cell**: raster cells extracted at the locations of the 'island shapefile' that are not complete cases and all adjacent cells that are not complete cases either. **Island shapefile**: the shapefile of Faial, Pico and São Jorge Islands.</span></div>


### cond.4.all()

To illustrate the function `cond.4.all()` and the other `scapesClassification` functions we will use a dummy raster ([Figure 2](#figure2)), compute its [attribute table](scapesClassification_01_1_FormatInputs.html#attribute-table), in which it is named `"dummy_var"` and compute its [list of neighborhoods](scapesClassification_01_1_FormatInputs.html#list-of-neighborhoods).

```{r def, message=FALSE, warning=FALSE, echo=TRUE}
# LOAD THE DUMMY RASTER
r <- list.files(system.file("extdata", package = "scapesClassification"), 
                 pattern = "dummy_raster\\.tif", full.names = T)
r <- raster(r)
```

```{r def_at, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# COMPUTE THE ATTRIBUTE TABLE
at <- attTbl(r, "dummy_var")

head(at)

# COMPUTE THE LIST OF NEIGBORHOODS
nbs <- ngbList(r)

head(nbs)
```

<a id="figure2"></a>

```{r dummy_rast_plot, message=FALSE, warning=FALSE, echo=TRUE}
# PLOT RASTER
par(mar = c(0.1, 5, 0.5, 0.1))
plot(r, axes=FALSE, box=FALSE, legend = FALSE)
text(r)
```

<div style="line-height: 0.8em; margin-bottom: 1.5em;"><span class="legend">**Figure 2 -** Dummy raster used to illustrate `scapesClassification` functions.</span></div>

---

The function `cond.4.all()` can be used to evaluate absolute test cell conditions (see `?conditions` and the article ["classificaton rules"](scapesClassification_01_3_ClassRules.html)) performing global evaluations (see the article ["rule evaluation"](scapesClassification_01_2_RuleEvaluation.html)). 

Using the dummy raster as an example, we can identify all cells having a value of 1. Classification rules are built accessing by name the variables stored in the attribute table (classification rule: `"dummy_var == 1"`). 

```{r cond_4_all1, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# ATTRIBUTE TABLE VARIABLES
names(at)

# COMPUTE THE CLASS VECTOR
cv   <- cond.4.all(attTbl = at, conditions = "dummy_var == 1", class = 10)

# CONVERT THE CLASS VECTOR INTO A RASTER
r_cv <- cv.2.rast(r, at$Cell, classVector = cv)
```

The element `cv` is a class vector. The classification number `10` (argument `class = 10`) was assigned to all cells having a value equal to 1 (argument `conditions = "dummy_var == 1"`). 

The element `r_cv` is the class vector `cv` converted into a raster by the function `cv.2.rast()`. It can be plotted and manipulated as a normal raster object ([Figure 3](#figure3)).

<a id="figure3"></a>
```{r cond_4_all1_plot, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# CLASS VECTOR LEVELS
unique(cv)

# class = 10, cell with values 'dummy_var == 1'
# class = NA, unclassified cells

# PLOT THE CLASS VECTOR
par(mar = c(0.1, 5, 0.5, 0.1))
plot(r_cv, axes=FALSE, box=FALSE, legend = FALSE, colNA="#818792", col="#78b2c4")
text(r)
```

<div style="line-height: 0.8em; margin-bottom: 1.5em;"><span class="legend">**Figure 3 -** Raster cell classification using the function **cond.4.all** and the rule `conditions = "dummy_var == 1"`. In **blue**, classified cells (corresponding to the positions `cv == 10`). In **grey**, unclassified cells.</span></div>

---

When no class vector is available, the function `cond.4.all()` can be used to begin a classification process, but it can also be used at other stages of a classification process. For instance, consider the rule `"dummy_var <= 3"`. This rule applied on our dummy raster would classify raster cells having values of 1, 2 and 3. However, providing the class vector `cv` as a function input, the function would update the class vector ignoring classified cells (non-NA cells in the class vector) ([Figure 4](#figure4)).

```{r cond_4_all2, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# UPDATE THE CLASS VECTOR

# UPDATED CLASS VECTOR
cv   <- cond.4.all(attTbl = at, conditions = "dummy_var <= 3", class = 1,
                   
                   # INITIAL CLASS VECTOR
                   classVector = cv) 

# CONVERT THE CLASS VECTOR INTO A RASTER
r_cv <- cv.2.rast(r, at$Cell, classVector = cv)
```

The class vector `cv` presents now two classification numbers, number `10` assigned in the previous evaluation and number `1` assigned in the last evaluation. 

<a id="figure4"></a>
```{r cond_4_all2_plot, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# CLASS VECTOR LEVELS
unique(cv)

# class = 10, cell with values 'dummy_var == 1'
# class = 1,  cell with values 'dummy_var =< 3'
# class = NA, unclassified cells

# PLOT THE CLASS VECTOR
par(mar = c(0.1, 2, 0.5, 0.1))
plot(r_cv, axes=FALSE, box=FALSE, legend = FALSE, colNA="#818792", col=c("#cfad89", "#78b2c4"))
text(r)
```

<div style="line-height: 0.8em; margin-bottom: 1.5em;"><span class="legend">**Figure 4 -** Two-step classification using the function **cond.4.all**. Step one (**light blue**), cells classified using the rule `conditions = "dummy_var == 1"`. Step two (**brown**), cells classified using the rule `conditions = "dummy_var <= 3"`. In **grey**, unclassified cells. Note that the rule of step two do not overwrite the classification of step one.</span></div>

## User condition functions

**Functions:** `cond.4.nofn()`, `cond.reclass()` and `cond.4.all()`.

User condition functions present the `conditions` argument and generally require a class vector as input. Excluding the function `cond.4.all()` (that can be considered both an anchor cell and a user condition function), user condition functions always perform [focal evaluations](scapesClassification_01_2_RuleEvaluation.html#focal-evaluation). Focal evaluations only test classification rules at positions adjacent to particular raster cells, anchor and focal cells. Anchor and focal cells are identified by the function argument `nofn`. 

A particular case of focal evaluation, is [class continuity](scapesClassification_01_2_RuleEvaluation.html#class-continuity). Class continuity join into the same class cells that share the same property(ies) and that are connected to the same anchor cell(s). User condition functions take into account class continuity when the classification number assigned with the argument `class` is also included in the argument `nofn`. 

### cond.4.nofn

The function `cond.4.nofn()` can be used to evaluate both relative and absolute conditions (see `?conditions` and ["classificaton rules"](scapesClassification_01_3_ClassRules.html)). We will start by presenting some absolute conditions. 

As all other user condition functions, `cond.4.nofn()` requires some focal cell class to be specified by the argument `nofn`. We can create an empty class vector and assume that `cell 25` is our initial focal cell. The focal cell will be identified by the numeric class `0`, but we could use any other number.

```{r cond_4_nofn0, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# SET CELL 32 AS A FOCAL CELL WITH CLASS = 10
at$cv[at$Cell == 25] <- 0
```

We can now use the function `cond.4.nofn()` to find cells with `"dummy_var >= 3"` at locations adjacent to the focal cell ([Figure 5](#figure5)).

```{r cond_4_nofn1, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# COMPUTE THE CLASS VECTOR (do not consider class continuity)
cv1 <- cond.4.nofn(attTbl = at, 
                   ngbList = nbs, 
                   
                   # CLASS VECTOR - INPUT
                   classVector = at$cv, 
                   
                   # FOCAL CELL CLASS
                   nbs_of = 0, 
                   
                   # CLASSIFICATION NUMBER
                   class = 1,  
                   
                   # ABSOLUTE CONDITION
                   conditions = "dummy_var >= 3")


# CONVERT THE CLASS VECTOR INTO A RASTER
r_cv1 <- cv.2.rast(r, at$Cell,classVector = cv1, plot = FALSE)
```

<a id="figure5"></a>
 
```{r cond_4_nofn1_plot, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# CLASS VECTOR LEVELS
unique(cv1)

# class = 0,  focal cell
# class = 1,  cell with values 'dummy_var >= 3'
# class = NA, unclassified cells

# PLOT THE CLASS VECTOR
par(mar = c(0.1, 2, 0.5, 0.1))
plot(r_cv1, axes=FALSE, box=FALSE, legend = FALSE, colNA="#818792", col=c("#78b2c4", "#cfad89"))
text(r)
```

<div style="line-height: 0.8em; margin-bottom: 1.5em;"><span class="legend">**Figure 5 -** Raster cell classification using the function **cond.4.nofn** and the rule `conditions = "dummy_var >= 3"`. In **light blue**, the focal cell; in **brown** cell meeting function conditions; in **grey**, unclassified cells. Only **cells contiguous** to the focal cell with values `>=3` are identified by the function.</span></div>

---

When the classification number associated with the argument `class` is also included in the argument `nofn`, classifications take into account [class continuity](scapesClassification_01_2_RuleEvaluation.html#class-continuity) ([Figure 6](#figure6)). 

```{r cond_4_nofn2, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# COMPUTE THE CLASS VECTOR (consider class continuity)
cv2 <- cond.4.nofn(attTbl = at, 
                   ngbList = nbs, 
                   
                   # CLASS VECTOR - INPUT
                   classVector = at$cv, 
                   
                   nbs_of = c(0,  # FOCAL CELL CLASS
                              1), # CLASSIFICATION NUMBER
                   
                   # CLASSIFICATION NUMBER 
                   class = 1,     # CLASSIFICATION NUMBER
                   
                   # ABSOLUTE CONDITION
                   conditions = "dummy_var >= 3")


# CONVERT THE CLASS VECTOR INTO A RASTER
r_cv2 <- cv.2.rast(r, at$Cell,classVector = cv2, plot = FALSE)
```

<a id="figure6"></a>

```{r cond_4_nofn2_plot, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# CLASS VECTOR LEVELS
unique(cv2)

# class = 0,  focal cell
# class = 1,  cell with values 'dummy_var >= 3'
# class = NA, unclassified cells

# PLOT THE CLASS VECTOR
par(mar = c(0.1, 2, 0.5, 0.1))
plot(r_cv2, axes=FALSE, box=FALSE, legend = FALSE, colNA="#818792", col=c("#78b2c4", "#cfad89"))
text(r)
```

<div style="line-height: 0.8em; margin-bottom: 1.5em;"><span class="legend">**Figure 6 -** Raster cell classification using the function **cond.4.nofn** and the rule `conditions = "dummy_var >= 3"`. In **light blue**, the focal cell; in **brown** cell meeting function conditions; in **grey**, unclassified cells. Only **cells contiguous AND continuous** to the focal cell with values `>=3` are identified by the function.</span></div>

---

Finally we will consider an [absolute neighborhood condition](scapesClassification_01_3_ClassRules.html#absolute-conditions). Absolute neighborhood conditions are flagged by a variable name followed by curly brackets (`{}`) and compare the values of the test cell and of its neighborhood against a threshold value. Test cells are classified if the rule is true for at least as many evaluations as the ones specified by the argument `fn_perc`. 

In the following example we will use the function `cond.4.nofn()` with the arguments `conditions = "dummy_var{} >= 3"` and  `fn_perc = 1`. This means that the function will identify ([Figure 7](#figure7)):

* cells whose value is `>= 3`;  

* _AND_ whose all (`fn_perc = 1`) neighbor values are `>= 3`;

* _AND_ that are contiguous and continuous to the focal cell.

```{r cond_4_nofn3, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# COMPUTE THE CLASS VECTOR (ABSOLUTE NEIGHBORHOOD CONDITION)
cv3 <- cond.4.nofn(attTbl = at, 
                   ngbList = nbs, 
                   
                   # CLASS VECTOR - INPUT
                   classVector = at$cv, 
                   
                   nbs_of = c(0,  # FOCAL CELL CLASS
                              1), # CLASSIFICATION NUMBER
                   
                   # CLASSIFICATION NUMBER 
                   class = 1,     # CLASSIFICATION NUMBER
                   
                   # ABSOLUTE NEIGHBORHOOD CONDITION
                   conditions = "dummy_var{} >= 3",
                   
                   # RULE HAS TO BE TRUE FOR CELL AND ALL ITS NEIGHBORHOOD
                   fn_perc = 1)


# CONVERT THE CLASS VECTOR INTO A RASTER
r_cv3 <- cv.2.rast(r, at$Cell,classVector = cv3, plot = FALSE)
```

<a id="figure7"></a>

```{r cond_4_nofn3_plot, message=FALSE, warning=FALSE, echo=TRUE, collapse=TRUE}
# CLASS VECTOR LEVELS
unique(cv3)

# class = 0,  focal cell
# class = 1,  cell with values 'dummy_var >= 3'
# class = NA, unclassified cells

# PLOT THE CLASS VECTOR
par(mar = c(0.1, 2, 0.5, 0.1))
plot(r_cv3, axes=FALSE, box=FALSE, legend = FALSE, colNA="#818792", col=c("#78b2c4", "#cfad89"))
text(r)
```

<div style="line-height: 0.8em; margin-bottom: 1.5em;"><span class="legend">**Figure 7 -** Raster cell classification using the function **cond.4.nofn**, the rule `conditions = "dummy_var{} >= 3"` and the argument `fn_perc = 1`. In **light blue**, the focal cell; in **brown** cell meeting function conditions; in **grey**, unclassified cells. Only cells with values `>=3`, with neighboring cells with values `>=3` and continuous with the focal cell are identified by the function.</span></div>

---

[Relative conditions](scapesClassification_01_3_ClassRules.html#relative-conditions) (conditions comparing the values of adjacent cells) can be evaluated by `cond.4.nofn`. 
